---
title: "PAA2024 - Extreme Heat and Cold Events Exposure Assessment with GridEX"
subtitle: "Connecting GridEX with PLACES health dataset"
author: "Vy Kim Nguyen"

format: html
editor: visual
---

## PAA2024 - Extreme Heat and Cold Events Exposure Assessment with GridEX

April 2024, PAA Annual Meeting

### Session 3B - Exploring the extreme events intensity and grid surfaces

Author: Vy Nguyen, Researcher at Harvard Medical School\
email: Vy_Nguyen\@hms.harvard.edu

PLACES provides estimates of health measures for counties, cities, neighborhoods, and ZIP code areas across the United States. These estimates currently cover 36 different chronic diseases.

<https://www.cdc.gov/places/about/index.html>

```{r}
library(sf)
library(dplyr)
library(data.table)

# workshop_dataset <- "../Data"

places_data_dir <- "../Data/PLACES_Dataset"
list.files(places_data_dir)
```

### Load the PLACES data

```{r}
places_county_2023 <- fread(file.path(places_data_dir, "PLACES__County_Data__GIS_Friendly_Format___2023_release.csv"))

class(places_county_2023)
dim(places_county_2023)
View(places_county_2023)
```

### Load and subset the extreme heat events frequency

```{r}
events_frequency_counties <- readRDS(file.path("../Data/GridEX_C15_T102_by_type_by_year_summaries","C15_S190_T102_events_by_type_by_year_Counties.rds"))

ehe_counties_ca_2022 <- events_frequency_counties %>% 
  dplyr::filter(event_type == "Extreme Heat Event") %>% 
  dplyr::filter(year_numerical == 2022)

class(ehe_counties_ca_2022)
dim(ehe_counties_ca_2022)
View(ehe_counties_ca_2022)
```

### Subset the PLACES data for state of california

```{r}
places_county_2023_ca <- places_county_2023 %>% 
  dplyr::filter(StateAbbr == "CA") %>% 
  mutate(CountyFIPS = paste0("0", as.character(CountyFIPS))) # adjust the GEOID code
View(places_county_2023_ca)
```

### Join the two dataset

```{r}
places_county_2023_ca_augmented <- places_county_2023_ca %>% 
  merge(.,
          ehe_counties_ca_2022,
          by.x = "CountyFIPS",
          by.y = "GEOID")
  
View(places_county_2023_ca_augmented)
```

### Explore the potential correlation

```{r}
library(tidyverse)

health_outcomes_codenames <- colnames(places_county_2023_ca) %>%
  .[grepl("_CrudePrev$|_AdjPrev", .)== TRUE]

correlation_ehe_health <- function(x
                                   , df_merged)
{
  print(x)
  
  health_outcome_i <- df_merged %>%
    pull(x)
  
  num_days_ehe <- df_merged %>%
    pull(total_event_days)
  
  correlation_i <- cor(health_outcome_i
                       , num_days_ehe)

  df_stats_i <- data.frame(health_outcome = x
                           , event_type = "Extreme Heat Event"
                           , correlation = correlation_i)
  return(df_stats_i)
}

df_correlations <- health_outcomes_codenames %>%
  map(.
      , correlation_ehe_health
      , places_county_2023_ca_augmented) %>%
  bind_rows()
View(df_correlations)
# cor(places_county_2023_ca_augmented$total_event_days, places_county_2023_ca_augmented$CASTHMA_AdjPrev)
```
