---
title: "Extreme Heat and Cold Events Exposure Assessment with GridEX"
subtitle: "Identifying vulnerable communities that experience extreme weather but lack adequate climate control resources in California"
author: "Pedram Fard, Researcher at Harvard Medical School"
email: pedram_fard@hms.harvard.edu
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: show
    theme: flatly
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, fig.width = 10, fig.height = 6)
```

## Overview
This analysis examines the intersection of extreme weather events (Extreme Cold Events - ECE and Extreme Heat Events - EHE) with housing infrastructure access to heating and cooling resources in California census tracts. The goal is to identify vulnerable communities that experience extreme weather but lack adequate climate control resources.


## Setup and Data Loading

```{r libraries}
# Load required libraries
library(tidycensus)
library(dplyr)
library(sf)
library(ggplot2)
library(knitr)
library(DT)
library(plotly)
library(purrr)
library(tidyr)
library(viridis)
library(scales)
library(RColorBrewer)

```

## Census Data Collection

```{r census-variables}
# Define heating and cooling variables from Census ACS
heating_cooling_vars <- c(
  # House heating fuel (Table B25040)
  "B25040_001",  # Total occupied housing units
  "B25040_002",  # Utility gas
  "B25040_003",  # Bottled, tank, or LP gas
  "B25040_004",  # Electricity
  "B25040_005",  # Fuel oil, kerosene, etc.
  "B25040_006",  # Coal or coke
  "B25040_007",  # Wood
  "B25040_008",  # Solar energy
  "B25040_009",  # Other fuel
  "B25040_010",  # No fuel used
  
  # Air conditioning (Table B25045)
  "B25045_001",  # Total occupied housing units (for AC)
  "B25045_002",  # With air conditioning
  "B25045_003"   # No air conditioning
)

# Function to get Census data for multiple states
get_census_data_for_states <- function(states, year = 2014) {
  map_dfr(states, function(state) {
    tryCatch({
      cat("Downloading data for state:", state, "\n")
      get_acs(
        geography = "tract",
        variables = heating_cooling_vars,
        state = state,
        year = year,
        survey = "acs5",
        geometry = FALSE,
        output = "wide"
      )
    }, error = function(e) {
      message(paste("Error downloading data for state", state, ":", e$message))
      return(NULL)
    })
  })
}
```


```{r download-census-data}
california_state_code <- "06"  # California FIPS code
census_heating_cooling <- get_census_data_for_states(california_state_code, year = 2014)
cat("Downloaded Census data for", nrow(census_heating_cooling), "census tracts\n")
```

## Data Processing and Cleaning

```{r clean-census-data}
# Clean and process Census data
census_clean <- census_heating_cooling %>%
  rename(
    total_housing_heating = B25040_001E,
    heating_utility_gas = B25040_002E,
    heating_bottled_gas = B25040_003E,
    heating_electricity = B25040_004E,
    heating_fuel_oil = B25040_005E,
    heating_coal = B25040_006E,
    heating_wood = B25040_007E,
    heating_solar = B25040_008E,
    heating_other = B25040_009E,
    heating_no_fuel = B25040_010E,
    total_housing_ac = B25045_001E,
    with_air_conditioning = B25045_002E,
    no_air_conditioning = B25045_003E
  ) %>%
  mutate(
    # Replace NA values with 0 for calculations
    across(starts_with("heating_"), ~ifelse(is.na(.x), 0, .x)),
    across(c(with_air_conditioning, no_air_conditioning), ~ifelse(is.na(.x), 0, .x)),
    
    # Calculate percentages and identify lack of resources
    pct_no_heating = ifelse(total_housing_heating > 0, 
                           (heating_no_fuel / total_housing_heating) * 100, 0),
    pct_no_ac = ifelse(total_housing_ac > 0, 
                      (no_air_conditioning / total_housing_ac) * 100, 0),
    
    # Define vulnerability thresholds (adjustable)
    lacks_heating = heating_no_fuel > 0.1,
    lacks_cooling = pct_no_ac > 0.1,
    
    
    total_with_heating = total_housing_heating - heating_no_fuel,
    pct_with_heating = ifelse(total_housing_heating > 0,
                             (total_with_heating / total_housing_heating) * 100, 0),
    
    # Primary heating source categories
    primary_heating_gas = heating_utility_gas + heating_bottled_gas,
    primary_heating_electric = heating_electricity,
    primary_heating_alternative = heating_wood + heating_coal + heating_solar + heating_other
  ) %>%
  select(GEOID, NAME, total_housing_heating, total_housing_ac,
         heating_utility_gas, heating_electricity, heating_wood, heating_no_fuel,
         primary_heating_gas, primary_heating_electric, primary_heating_alternative,
         with_air_conditioning, no_air_conditioning,
         pct_no_heating, pct_no_ac, pct_with_heating,
         lacks_heating, lacks_cooling)

cat("Processed", nrow(census_clean), "census tracts with heating/cooling data\n")
```
# Join with the GridEX data
```{r join-data}
# Filter climate data for California and join with Census data
Y2014_S201_T002_impacted_area_variables_intensity_tracts <- readRDS(url("https://dataverse.harvard.edu/api/access/datafile/10590009"))

california_climate <- Y2014_S201_T002_impacted_area_variables_intensity_tracts %>%
  filter(substr(GEOID, 1, 2) == "06") %>% 
  mutate(
    # Clean event type classification
    event_type_clean = case_when(
      grepl("Cold", event_type, ignore.case = TRUE) ~ "ECE",
      grepl("Heat", event_type, ignore.case = TRUE) ~ "EHE",
      TRUE ~ as.character(event_type)
    )
  )

glimpse(california_climate)
#View(california_climate)

# Join climate data with Census data
combined_data <- california_climate %>%
  inner_join(census_clean, by = "GEOID") %>%
  mutate(
    has_ece = event_type_clean == "ECE",
    has_ehe = event_type_clean == "EHE",
    
    # Create vulnerability flags
    cold_vulnerable = has_ece & lacks_heating,
    heat_vulnerable = has_ehe & lacks_cooling,
    
    # Overall vulnerability
    climate_vulnerable = cold_vulnerable | heat_vulnerable,
    
    # Extract county from NAME for grouping
    #county = gsub(".*,\\s*(.*)\\s*County.*", "\\1", NAME.y)
    county = trimws(gsub(".*,\\s*(.*)\\s*County.*", "\\1", NAME.y))
  )

glimpse(combined_data)
#View(combined_data)

cat("Successfully joined data for", nrow(combined_data), "California census tracts\n")
```

## Get Additional Geographic Data for Mapping

```{r get-geography}
# Get California census tract geometry
ca_tracts_geo <- get_acs(
  geography = "tract",
  variables = "B01003_001",  # Total population for reference
  state = "06",  
  year = 2014,
  survey = "acs5",
  geometry = TRUE
) %>%
  select(GEOID, geometry)

# Get California county boundaries for reference
ca_counties_geo <- get_acs(
  geography = "county",
  variables = "B01003_001",
  state = "06",
  year = 2014,
  survey = "acs5",
  geometry = TRUE
) %>%
  select(GEOID, NAME, geometry)

cat("Downloaded geographic data for", nrow(ca_tracts_geo), "census tracts\n")
cat("Downloaded geographic data for", nrow(ca_counties_geo), "counties\n")
```

```{r create-spatial-data}
# Join geographic data with vulnerability analysis
vulnerability_map_data <- ca_tracts_geo %>%
  left_join(combined_data, by = "GEOID") %>%
  mutate(
    # Create vulnerability categories for mapping
    vulnerability_category = case_when(
      cold_vulnerable & heat_vulnerable ~ "Both Cold & Heat Vulnerable",
      cold_vulnerable ~ "Cold Vulnerable",
      heat_vulnerable ~ "Heat Vulnerable",
      has_ece | has_ehe ~ "Has Extreme Events (Not Vulnerable)",
      TRUE ~ "No Extreme Events"
    ),
    
    # Create numeric vulnerability score for continuous mapping
    vulnerability_score = case_when(
      cold_vulnerable & heat_vulnerable ~ 3,
      cold_vulnerable | heat_vulnerable ~ 2,
      has_ece | has_ehe ~ 1,
      TRUE ~ 0
    ),
    
    # Calculate combined vulnerability percentage
    combined_vulnerability_pct = case_when(
      cold_vulnerable ~ pct_no_heating,
      heat_vulnerable ~ pct_no_ac,
      cold_vulnerable & heat_vulnerable ~ (pct_no_heating + pct_no_ac) / 2,
      TRUE ~ 0
    )
  )

# Filter to only tracts with data
vulnerability_map_data <- vulnerability_map_data %>%
  filter(!is.na(event_type_clean))

cat("Created spatial data for", nrow(vulnerability_map_data), "census tracts with vulnerability data\n")
```

## Vulnerability Analysis

### Summary Statistics

```{r vulnerability-summary}
# Overall vulnerability summary
vulnerability_summary <- combined_data %>%
  summarise(
    total_tracts = n(),
    tracts_with_ece = sum(has_ece, na.rm = TRUE),
    tracts_with_ehe = sum(has_ehe, na.rm = TRUE),
    ece_lacking_heating = sum(cold_vulnerable, na.rm = TRUE),
    ehe_lacking_cooling = sum(heat_vulnerable, na.rm = TRUE),
    total_vulnerable = sum(climate_vulnerable, na.rm = TRUE),
    avg_pct_no_heating = round(mean(pct_no_heating, na.rm = TRUE), 2),
    avg_pct_no_ac = round(mean(pct_no_ac, na.rm = TRUE), 2),
    .groups = 'drop'
  )

kable(vulnerability_summary, 
      caption = "California Climate Vulnerability Summary",
      col.names = c("Total Tracts", "ECE Tracts", "EHE Tracts", 
                   "Cold Vulnerable", "Heat Vulnerable", "Total Vulnerable",
                   "Avg % No Heating", "Avg % No AC"))
```

### County-Level Analysis

```{r county-analysis}
# County-level vulnerability analysis
county_analysis <- combined_data %>%
  group_by(county) %>%
  summarise(
    total_tracts = n(),
    ece_tracts = sum(has_ece, na.rm = TRUE),
    ehe_tracts = sum(has_ehe, na.rm = TRUE),
    cold_vulnerable_tracts = sum(cold_vulnerable, na.rm = TRUE),
    heat_vulnerable_tracts = sum(heat_vulnerable, na.rm = TRUE),
    total_vulnerable_tracts = sum(climate_vulnerable, na.rm = TRUE),
    vulnerability_rate = round((total_vulnerable_tracts / total_tracts) * 100, 1),
    avg_pct_no_heating = round(mean(pct_no_heating, na.rm = TRUE), 2),
    avg_pct_no_ac = round(mean(pct_no_ac, na.rm = TRUE), 2),
    .groups = 'drop'
  ) %>%
  arrange(desc(total_vulnerable_tracts))

# Display top 15 counties
DT::datatable(county_analysis, 
              caption = "County-Level Climate Vulnerability Analysis",
              options = list(pageLength = 15, scrollX = TRUE)) %>%
  DT::formatStyle("vulnerability_rate",
                  background = DT::styleColorBar(county_analysis$vulnerability_rate, 'lightblue'),
                  backgroundSize = '100% 90%',
                  backgroundRepeat = 'no-repeat',
                  backgroundPosition = 'center')
```

## Detailed Vulnerability Results

### Cold-Vulnerable Census Tracts

```{r cold-vulnerable}
# Census tracts with ECE that lack heating resources
ece_no_heating <- combined_data %>%
  filter(cold_vulnerable) %>%
  select(GEOID, NAME.y, county, event_type_clean, 
         pct_no_heating, heating_no_fuel, total_housing_heating,
         d_apparent_temperature_avg_min_intensity, 
         lacks_heating) %>%
  arrange(desc(pct_no_heating)) %>%
  mutate(
    pct_no_heating = round(pct_no_heating, 2),
    temp_intensity = round(d_apparent_temperature_avg_min_intensity, 1)
  )

if(nrow(ece_no_heating) > 0) {
  cat("Found", nrow(ece_no_heating), "census tracts with Extreme Cold Events lacking adequate heating\n\n")
  
  DT::datatable(ece_no_heating, 
                caption = "Census Tracts with ECE Lacking Adequate Heating",
                colnames = c("GEOID", "Name", "County", "Event Type", 
                           "% No Heating", "Units No Fuel", "Total Units",
                           "Min Temp Intensity", "Lacks Heating"),
                options = list(pageLength = 10, scrollX = TRUE))
} else {
  cat("No census tracts found with ECE lacking adequate heating in California\n")
}
```

### Heat-Vulnerable Census Tracts

```{r heat-vulnerable}
# Census tracts with EHE that lack cooling resources
ehe_no_cooling <- combined_data %>%
  filter(heat_vulnerable) %>%
  select(GEOID, NAME.y, county, event_type_clean,
         pct_no_ac, no_air_conditioning, total_housing_ac,
         d_apparent_temperature_avg_max_intensity,
         lacks_cooling) %>%
  arrange(desc(pct_no_ac)) %>%
  mutate(
    pct_no_ac = round(pct_no_ac, 2),
    temp_intensity = round(d_apparent_temperature_avg_max_intensity, 1)
  )

if(nrow(ehe_no_cooling) > 0) {
  cat("Found", nrow(ehe_no_cooling), "census tracts with Extreme Heat Events lacking adequate cooling\n\n")
  
  DT::datatable(ehe_no_cooling, 
                caption = "Census Tracts with EHE Lacking Adequate Cooling",
                colnames = c("GEOID", "Name", "County", "Event Type",
                           "% No AC", "Units No AC", "Total Units",
                           "Max Temp Intensity", "Lacks Cooling"),
                options = list(pageLength = 10, scrollX = TRUE))
} else {
  cat("No census tracts found with EHE lacking adequate cooling in California\n")
}
```


## Visualizations

### Most Climate-Vulnerable Counties - Heat Events

```{r county-heat-heatmap, fig.width=12, fig.height=8}
# County vulnerability heatmap for Heat Events
county_heat_data <- combined_data %>%
  filter(has_ehe) %>%
  group_by(county) %>%
  summarise(
    total_tracts = n(),
    heat_vulnerable_tracts = sum(heat_vulnerable, na.rm = TRUE),
    vulnerability_rate = round((heat_vulnerable_tracts / total_tracts) * 100, 1),
    avg_pct_no_ac = round(mean(pct_no_ac, na.rm = TRUE), 2),
    avg_temp_intensity = round(mean(d_apparent_temperature_avg_max_intensity, na.rm = TRUE), 1),
    .groups = 'drop'
  ) %>%
  filter(total_tracts >= 3) %>%  # Filter counties with at least 3 tracts
  top_n(15, heat_vulnerable_tracts)

p_heat <- ggplot(county_heat_data, aes(x = reorder(county, heat_vulnerable_tracts), 
                                       y = heat_vulnerable_tracts)) +
  geom_col(aes(fill = vulnerability_rate), width = 0.7) +
  coord_flip() +
  scale_fill_gradient2(low = "lightyellow", mid = "orange", high = "red",
                       midpoint = median(county_heat_data$vulnerability_rate),
                       name = "Vulnerability\nRate (%)") +
  labs(
    title = "Most Heat-Vulnerable Counties in California",
    subtitle = "Counties with Extreme Heat Events - vulnerable census tracts by vulnerability rate",
    x = "County",
    y = "Number of Heat-Vulnerable Census Tracts",
    caption = "Only counties with ≥3 census tracts with EHE shown"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 14, face = "bold"),
    plot.subtitle = element_text(size = 12),
    axis.text = element_text(size = 10)
  )

print(p_heat)
```

### Most Climate-Vulnerable Counties - Cold Events

```{r county-cold-heatmap, fig.width=12, fig.height=8}
# County vulnerability heatmap for Cold Events
county_cold_data <- combined_data %>%
  filter(has_ece) %>%
  group_by(county) %>%
  summarise(
    total_tracts = n(),
    cold_vulnerable_tracts = sum(cold_vulnerable, na.rm = TRUE),
    vulnerability_rate = round((cold_vulnerable_tracts / total_tracts) * 100, 1),
    avg_pct_no_heating = round(mean(pct_no_heating, na.rm = TRUE), 2),
    avg_temp_intensity = round(mean(d_apparent_temperature_avg_min_intensity, na.rm = TRUE), 1),
    .groups = 'drop'
  ) %>%
  filter(total_tracts >= 3) %>%  # Filter counties with at least 3 tracts
  top_n(15, cold_vulnerable_tracts)

p_cold <- ggplot(county_cold_data, aes(x = reorder(county, cold_vulnerable_tracts), 
                                       y = cold_vulnerable_tracts)) +
  geom_col(aes(fill = vulnerability_rate), width = 0.7) +
  coord_flip() +
  scale_fill_gradient2(low = "lightblue", mid = "purple", high = "blue",
                       midpoint = median(county_cold_data$vulnerability_rate),
                       name = "Vulnerability\nRate (%)") +
  labs(
    title = "Most Cold-Vulnerable Counties in California",
    subtitle = "Counties with Extreme Cold Events - vulnerable census tracts by vulnerability rate",
    x = "County",
    y = "Number of Cold-Vulnerable Census Tracts",
    caption = "Only counties with ≥3 census tracts with ECE shown"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 14, face = "bold"),
    plot.subtitle = element_text(size = 12),
    axis.text = element_text(size = 10)
  )

print(p_cold)
```

## Vulnerability Maps
### Heat-Vulnerable Counties Map
```{r heat-vulnerable-counties-map, fig.width=14, fig.height=12}
# Create county-level heat vulnerability data
county_heat_vulnerability <- combined_data %>%
  filter(has_ehe) %>% 
  group_by(county) %>%
  summarise(
    total_tracts = n(),
    heat_vulnerable_tracts = sum(heat_vulnerable, na.rm = TRUE),
    vulnerability_rate = round((heat_vulnerable_tracts / total_tracts) * 100, 1),
    avg_pct_no_ac = round(mean(pct_no_ac, na.rm = TRUE), 2),
    .groups = 'drop'
  ) %>%
  mutate(
    vulnerability_level = case_when(
      vulnerability_rate >= 75 ~ "Very High (≥75%)",
      vulnerability_rate >= 50 ~ "High (50-74%)",
      vulnerability_rate >= 25 ~ "Moderate (25-49%)",
      vulnerability_rate >= 10 ~ "Low (10-24%)",
      vulnerability_rate > 0 ~ "Very Low (1-9%)",
      TRUE ~ "No Vulnerability"
    ),
    # Create factor with levels in the order they appear in case_when
    vulnerability_level = factor(vulnerability_level, levels = c(
      "Very High (≥75%)",
      "High (50-74%)", 
      "Moderate (25-49%)",
      "Low (10-24%)",
      "Very Low (1-9%)",
      "No Vulnerability",
      "No EHE Data"
    ))
  ) 

# Join with county geographic data - keep all counties
heat_counties_map <- ca_counties_geo %>%
  mutate(county = trimws(gsub("\\s*County.*", "", NAME))) %>%
  left_join(county_heat_vulnerability, by = "county") %>%
  mutate(
    vulnerability_level = ifelse(is.na(vulnerability_level), "No EHE Data", as.character(vulnerability_level)),
    vulnerability_level = factor(vulnerability_level, levels = c(
      "Very High (≥75%)",
      "High (50-74%)", 
      "Moderate (25-49%)",
      "Low (10-24%)",
      "Very Low (1-9%)",
      "No Vulnerability",
      "No EHE Data"
    ))
  )

glimpse(county_heat_vulnerability)

# Heat vulnerability color palette for counties (reordered to match factor levels)
heat_county_colors <- c(
  "Very High (≥75%)" = "#800026",
  "High (50-74%)" = "#e31a1c",
  "Moderate (25-49%)" = "#fd8d3c", 
  "Low (10-24%)" = "#fed976",
  "Very Low (1-9%)" = "#ffffcc",
  "No Vulnerability" = "#90FF50",
  "No EHE Data" = "#f0f0f0"
)

heat_counties_vulnerability_map <- ggplot() +
  # Add heat vulnerability data for all counties
  geom_sf(data = heat_counties_map,
          aes(fill = vulnerability_level),
          color = "white", size = 0.5) +
  scale_fill_manual(values = heat_county_colors,
                    name = "Heat Vulnerability Level",
                    guide = guide_legend(override.aes = list(size = 0.5))) +
  labs(
    title = "Heat-Vulnerable Counties in California",
    subtitle = "Counties with Extreme Heat Events - % of census tracts lacking adequate cooling (any % no AC)",
  ) +
  theme_void() +
  theme(
    plot.title = element_text(size = 16, face = "bold", hjust = 0.5),
    plot.subtitle = element_text(size = 12, hjust = 0.5, margin = margin(b = 20)),
    plot.caption = element_text(size = 10, hjust = 0.5, margin = margin(t = 20)),
    legend.position = "bottom",
    legend.title = element_text(size = 11, face = "bold"),
    legend.text = element_text(size = 10),
    legend.key.size = unit(0.8, "cm"),
    plot.margin = margin(20, 20, 20, 20)
  )

print(heat_counties_vulnerability_map)
```

### Cold-Vulnerable Counties Map
```{r cold-vulnerable-counties-map, fig.width=14, fig.height=12}
# Create county-level cold vulnerability data
county_cold_vulnerability <- combined_data %>%
  filter(has_ece) %>%
  group_by(county) %>%
  summarise(
    total_tracts = n(),
    cold_vulnerable_tracts = sum(cold_vulnerable, na.rm = TRUE),
    vulnerability_rate = round((cold_vulnerable_tracts / total_tracts) * 100, 1),
    avg_pct_no_heating = round(mean(pct_no_heating, na.rm = TRUE), 2),
    .groups = 'drop'
  ) %>%
  mutate(
    vulnerability_level = case_when(
      vulnerability_rate >= 75 ~ "Very High (≥75%)",
      vulnerability_rate >= 50 ~ "High (50-74%)",
      vulnerability_rate >= 25 ~ "Moderate (25-49%)",
      vulnerability_rate >= 10 ~ "Low (10-24%)",
      vulnerability_rate > 0 ~ "Very Low (1-9%)",
      TRUE ~ "No Vulnerability"
    ),
    # Create factor with levels in the order they appear in case_when
    vulnerability_level = factor(vulnerability_level, levels = c(
      "Very High (≥75%)",
      "High (50-74%)", 
      "Moderate (25-49%)",
      "Low (10-24%)",
      "Very Low (1-9%)",
      "No Vulnerability",
      "No ECE Data"
    ))
  )

# Join with county geographic data - keep all counties
cold_counties_map <- ca_counties_geo %>%
  mutate(county = trimws(gsub("\\s*County.*", "", NAME))) %>%
  left_join(county_cold_vulnerability, by = "county") %>%
  mutate(
    vulnerability_level = ifelse(is.na(vulnerability_level), "No ECE Data", as.character(vulnerability_level)),
    vulnerability_level = factor(vulnerability_level, levels = c(
      "Very High (≥75%)",
      "High (50-74%)", 
      "Moderate (25-49%)",
      "Low (10-24%)",
      "Very Low (1-9%)",
      "No Vulnerability",
      "No ECE Data"
    ))
  )

# Cold vulnerability color palette for counties (reordered to match factor levels)
cold_county_colors <- c(
  "Very High (≥75%)" = "#08306b",
  "High (50-74%)" = "#2171b5", 
  "Moderate (25-49%)" = "#6baed6",
  "Low (10-24%)" = "#c6dbef",
  "Very Low (1-9%)" = "#f7fbff",
  "No Vulnerability" = "#90FF50",
  "No ECE Data" = "#f0f0f0"
)

cold_counties_vulnerability_map <- ggplot() +
  # Add cold vulnerability data for all counties
  geom_sf(data = cold_counties_map,
          aes(fill = vulnerability_level),
          color = "white", size = 0.5) +
  scale_fill_manual(values = cold_county_colors,
                    name = "Cold Vulnerability Level",
                    guide = guide_legend(override.aes = list(size = 0.5))) +
  labs(
    title = "Cold-Vulnerable Counties in California", 
    subtitle = "Counties with Extreme Cold Events - % of census tracts lacking heating (any units without fuel)",
  ) +
  theme_void() +
  theme(
    plot.title = element_text(size = 16, face = "bold", hjust = 0.5),
    plot.subtitle = element_text(size = 12, hjust = 0.5, margin = margin(b = 20)),
    plot.caption = element_text(size = 10, hjust = 0.5, margin = margin(t = 20)),
    legend.position = "bottom",
    legend.title = element_text(size = 11, face = "bold"),
    legend.text = element_text(size = 10),
    legend.key.size = unit(0.8, "cm"),
    plot.margin = margin(20, 20, 20, 20)
  )

print(cold_counties_vulnerability_map)
```