---
title: "Exercise"
output: html_document
date: '2024-04-15'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Based on the code used during the previous sessions, please complete the following tasks:

1. Retrieve the county boundaries for a state of your choice using the tigris package;

2. Load the Gridex data in a raster format (for 2022 or another year) and extract the average values for heat for each county;

3. Retrieve data on the percentage of individuals aged above 65 in the counties in the state of your choice;

4. Merge the two datasets and provide some summary statistics for the number of heat events and the percentage of inviduals aged above 65.




##Solutions

```{r ,  message=FALSE, warning=FALSE, error=FALSE}
# Library
rm(list=ls(all=TRUE))

library(sf)
library(tidyverse)
library(ggplot2)
library(mapview)
library(terra)
library(tigris)
library(tidycensus)
library(exactextractr)
library(purrr)

```

### Step 1.

```{r ,  message=FALSE, warning=FALSE, error=FALSE}
counties_ny = counties("NY")
mapview(counties_ny)
```


### Step 2. 

```{r ,  message=FALSE, warning=FALSE, error=FALSE }
ehe = rast("C:/Users/contekeivabu/Desktop/TEACHING/PAA_2024/data/Y2022_S426_T102_ehe_unified_grids_500_cumulative.tif")
```


```{r ,  message=FALSE, warning=FALSE, error=FALSE}
(
counties_ny_crs <- st_transform(counties_ny, st_crs(ehe))
)
ehe_crop <- terra::crop(ehe, counties_ny_crs)
plot(ehe_crop)
###extract
counties_ny_crs$heat <- exact_extract(ehe_crop, counties_ny_crs, 'mean' , progress = FALSE)
head(counties_ny_crs)
```

### Step 3. 
```{r ,  message=FALSE, warning=FALSE, error=FALSE}
vars <- c("B01001_001",  # Total population
          "B01001_020",  # Male: 65 and 66 years
          "B01001_021",  # Male: 67 to 69 years
          "B01001_022",  # Male: 70 to 74 years
          "B01001_023",  # Male: 75 to 79 years
          "B01001_024",  # Male: 80 to 84 years
          "B01001_025",  # Male: 85 years and over
          "B01001_044",  # Female: 65 and 66 years
          "B01001_045",  # Female: 67 to 69 years
          "B01001_046",  # Female: 70 to 74 years
          "B01001_047",  # Female: 75 to 79 years
          "B01001_048",  # Female: 80 to 84 years
          "B01001_049")  # Female: 85 years and over



# Fetch data for a specific census tract in a given state and county
data_age <- get_acs(geography = "county", 
                variables = vars, 
                state = "NY" ,
                survey = "acs5",
                geometry= FALSE,
                year = 2021)
###
county_data_age <- data_age %>%
  group_by(GEOID) %>%
  summarize(total_pop = sum(estimate[variable == "B01001_001"]),
            age_pop = sum(estimate[variable %in% vars[-1]])) %>%
  mutate(proportion_age = 100 * (age_pop / total_pop)) %>%
  select(GEOID, total_pop, age_pop, proportion_age)
head(county_data_age)

```


### Step 4. 

```{r ,  message=FALSE, warning=FALSE, error=FALSE}
merged_age_heat <- merge(county_data_age, counties_ny_crs, by = "GEOID")
##
ggplot(merged_age_heat, aes(x = proportion_age, y = heat)) +
  geom_point() +
  geom_smooth(method = "loess", se = FALSE, color = "blue") +
  labs(title = "Percentage above age 65 and Extreme Heat in New York",
       x = "Percentage > age 65",
       y = "Total Event Heat Days 2022") +
  theme_minimal()
```



