---
title: "PAA2024 - Extreme Heat and Cold Events Exposure Assessment with GridEX"
subtitle: "Exploring the aggregate impacts of extreme events"
author: "Pedram Fard, Researcher at Harvard Medical School"
email: pedram_fard@hms.harvard.edu

format: html
editor: visual
---

email: pedram_fard\@hms.harvard.edu

### April 2024, PAA Annual Meeting

## Session 2 - Part C - Exploring the impacts of extreme events within administrative boundaries

In this notebook we explore techniques to delineate spatial extent of extreme events using interpolated surfaces.

```{r, message=TRUE, warning=FALSE, error=TRUE}
library(terra)
library(sf)
library(dplyr)

#workshop_dataset <- "../Data"
# Check out the data folders for today workshop
#list.files(workshop_dataset, full.names = TRUE)

# Load the spatial layers of administrative boundaries 
census_data_dir <- "../Data/Census_admin_boundaries"

us_states <- st_read(file.path(census_data_dir, "tigris_states_geo_CRS5070.gpkg"))
us_counties <- st_read(file.path(census_data_dir, "tigris_counties_geo_CRS5070.gpkg"))

daily_data_dir <- "../Data/GridEX_T002_daily_raster_dataset"
list.files(daily_data_dir, pattern = "idw")
```

```{r}
ehce_daily_raster_22Jan03 <- rast(file.path(daily_data_dir, "D2022-01-03_S526_T002_idw_ehce_grid.tif"))

plot(ehce_daily_raster_22Jan03)
plot(us_states[2], add = T, col = NA) 

```

### Processing the events intensity surfaces using raster operations

```{r}
ece_daily_raster_22Jan03 <- ehce_daily_raster_22Jan03
ece_daily_raster_22Jan03[(ece_daily_raster_22Jan03 < 0)] <- -1 
ece_daily_raster_22Jan03[(ece_daily_raster_22Jan03 > 0)] <- NA

plot(ece_daily_raster_22Jan03, col = "blue")
plot(us_states[2], add = T, col = NA)
```

Loading the same layer from the pre-computed dataset

```{r}
# ece_daily_raster_22Jan03 <- rast(file.path(daily_data_dir, "D2022-01-03_S422_T002_idw_ece_dissolved_grid.tif"))
# plot(ece_daily_raster_22Jan03, col = "blue")
# plot(us_states[2], add = T, col = NA)
```

### Identifying a sample event boundaries

Delineate the vector boundaries through vectorizing the dissolved raster layer using methods from "terra" library.

```{r}
idw_ehce_vector_boundaries_22Jan03 <- 
as.polygons(ece_daily_raster_22Jan03, dissolve=TRUE) %>% 
fillHoles(.)

dim(idw_ehce_vector_boundaries_22Jan03)
#View(idw_ehce_vector_boundaries_22Jan03)

```

```{r}
plot(idw_ehce_vector_boundaries_22Jan03)
plot(us_states[2], add = T, col = NA)

```

Converting the primarily derived vector into a more flexible format using methods from "sf" library.

```{r}
idw_ehce_boundaries_sf_22Jan03 <- idw_ehce_vector_boundaries_22Jan03[""] %>% 
    st_as_sf() %>% # Convert SpatVector data to 'sf' object
    st_make_valid() %>% # Resolve potential topological errors
    st_cast("POLYGON") %>%  # Resolve potential topological errors
    #st_transform() %>% 
    #st_cast("MULTIPOLYGON") %>% 
    mutate(event_type = "ECE") %>%
    mutate(event_date = as.Date("01-03-2022", "%m-%d-%Y")) %>%
    mutate(ehce_geo_id = paste0(sprintf("%03d", row_number()))) %>% 
    
    mutate(computed_event_area = round(as.numeric(st_area(.)), 0)) %>%
    mutate(area_hectare = round((computed_event_area / 10000), 1)) 

class(idw_ehce_boundaries_sf_22Jan03)
glimpse(idw_ehce_boundaries_sf_22Jan03)

```

Read the yearly dataset of pre-computed EHE/ECE boundaries

```{r}
daily_vectors_dir <- "../Data/GridEX_S123_T002_daily_vector_dataset_GPKG"
list.files(daily_vectors_dir, "vector_boundaries")

ehe_ece_boundaries_2022 <- st_read(file.path(daily_vectors_dir, "Y2022_S400_T002_ehe_ece_vector_boundaries_CRS5070.gpkg"))

#print(ehe_ece_boundaries_2022)
#View(ehe_ece_boundaries_2022)

#head(ehe_ece_boundaries_2022)
ls(ehe_ece_boundaries_2022)
```

### Mapping the combination of raster and vector layers for a sample day of ECE

Load the pre-computed raster layers, and subset the data for a specific day of ECE (Jan 03, 2022)

```{r}
ehe_ece_boundaries_2022_01_03 <- ehe_ece_boundaries_2022 %>% 
  dplyr::filter(event_date == "2022-01-03")

plot(ece_daily_raster_22Jan03, col = "blue")
plot(ehe_ece_boundaries_2022_01_03["area_hectare"], col = NA, add = T)

#plot(ehce_daily_raster_22Jan03)
#plot(ehe_ece_boundaries_2022_01_03["area_hectare"], col = NA, add = T)
```

Load the additional variables computed for the impacted area

```{r}
ece_daily_raster_vars_stacks_22Jan03 <- rast(file.path(daily_data_dir, "D2022-01-03_daily_idw_variables_stack.tif"))

plot(ece_daily_raster_vars_stacks_22Jan03)
```

```{r}
plot(ece_daily_raster_vars_stacks_22Jan03["EHCMI_normalized"])
```

## Inspection of EHE/ECE overlayed with administrative boundaries

```{r}
plot(ehce_daily_raster_22Jan03)
#plot(ehe_ece_boundaries_2022_01_03["area_hectare"], col = NA)
plot(us_counties[2], col = NA, add = T)

```

In Session 3, we will explore the process of extracting values from grid layers into vector layers with more details.
