---
title: "PAA2024 - Extreme Heat and Cold Events Exposure Assessment with GridEX"
subtitle: "Exploring the aggregate impacts of extreme events"
author: "Pedram Fard, Researcher at Harvard Medical School"
email: pedram_fard@hms.harvard.edu

format: html
editor: visual
---

email: pedram_fard\@hms.harvard.edu

### April 2024, PAA Annual Meeting

## Session 2 - Part C - Exploring the daily impacts of extreme events within administrative boundaries

In this notebook we explore the disaggregate daily vector layers representing the spatial extent of extreme events under administrative boundaries.

GrideEX uses FIPS code to identify Census and Administrative boundaries. For a look up reference of FIPS codes please see:\
[Federal Information Processing System (FIPS) Codes for States and Counties](https://transition.fcc.gov/oet/info/maps/census/fips/fips.txt)

```{r, message=TRUE, warning=FALSE, error=TRUE}
library(terra)
library(sf)

#workshop_dataset <- "../Data"
# Check out the data folders for today workshop
#list.files(workshop_dataset, full.names = TRUE)

# Load the spatial layers of administrative boundaries 
census_data_dir <- "../Data/Census_admin_boundaries"

us_states <- st_read(file.path(census_data_dir, "tigris_states_geo_CRS5070.gpkg"))
us_counties <- st_read(file.path(census_data_dir, "tigris_counties_geo_CRS5070.gpkg"))

daily_vectors_dir <- "../Data/GridEX_S123_T002_daily_vector_dataset_GPKG"
list.files(daily_vectors_dir)
```

"S190" indicates that we are loading the County level data. "T002" means this file only includes records of the days experiencing at least one type of extreme events.

```{r}
impacted_areas_CA_2022 <- st_read(file.path(daily_vectors_dir, "Y2022_S190_T002_impacted_area_variables_intensity_Counties_wideformat_CRS4269_CA.gpkg"))

#View(impacted_areas_CA_2022)
ls(impacted_areas_CA_2022)

# number of events recorded in California on 2022
length(unique(impacted_areas_CA_2022$event_date))
```

```{r}
#head(impacted_areas_CA_2022)

impacted_areas_CA_2022_01_03 <- impacted_areas_CA_2022 %>% 
  dplyr::filter(event_date == "2022-01-03")

plot(impacted_areas_CA_2022["ehcmi_normalized_median_intensity"])
```

When analyzing the daily impacts data particularly for extended duration, it is recommended to work with tabular data rather than spatial objects. After performing your analysis you can then join the final results to the associated geometries.

```{r}
list.files(daily_vectors_dir)

C15_counties_impacted_areas <- readRDS(file.path(daily_vectors_dir, "C15_S190_T002_impacted_vars_intensity_Counties_TB.rds"))

mem_size <- scales::label_bytes()(object.size(C15_counties_impacted_areas)[1])
cat("Size of object in memory:",  mem_size)

head(C15_counties_impacted_areas)
```
