---
title: "PAA2024 - Extreme Heat and Cold Events Exposure Assessment with GridEX"
subtitle: "Exploring the cumulative impacts of extreme events"
author: "Pedram Fard, Researcher at Harvard Medical School"
email: pedram_fard@hms.harvard.edu

format: html
editor: visual
---

email: pedram_fard\@hms.harvard.edu

### April 2024, PAA Annual Meeting

## Session 2 - Part D - Exploring the cumulative impacts of extreme events

In this notebook we work with the vector layers representing the cumulative impacts of extreme heat and cold events (EHE/ECE) in the US.

For more information see: [Spatio-temporal interpolation and delineation of extreme heat events in California between 2017 and 2021](https://www.sciencedirect.com/science/article/pii/S0013935123017887?dgcid=rss_sd_all)

```{r, message=TRUE, warning=FALSE, error=TRUE}

library(sf)
library(dplyr)
library(leaflet)

#workshop_dataset <- "../Data"
## Check out the data folders for today workshop
#list.files(workshop_dataset, full.names = TRUE)

# Load the spatial layers of administrative boundaries 
census_data_dir <- "../Data/Census_admin_boundaries"

us_states <- st_read(file.path(census_data_dir, "tigris_states_geo_CRS5070.gpkg"))
us_counties <- st_read(file.path(census_data_dir, "tigris_counties_geo_CRS5070.gpkg"))
```

"S190" indicates that we are loading the County level data. "T102" means this file contains aggregate yearly frequency of extreme event days.

```{r}
events_frequency_counties <- readRDS(file.path("../Data/GridEX_C15_T102_by_type_by_year_summaries", "C15_S190_T102_events_by_type_by_year_sf_Counties.rds"))

glimpse(events_frequency_counties)

```

```{r}
# Function to create a Leaflet map
create_leaflet_map <- function(data, event_type = "Extreme Heat Event") {
  pal <- colorNumeric(
    palette = if (event_type == "Extreme Heat Event") 
      c("white", "orange", "red3", "brown4")
    else 
      c("white", "cyan4", "blue", "purple4"),
    domain = data$total_event_days
  )
  
  leaflet(data) %>%
    setView(lng = -98.5, lat = 39.8, zoom = 4) %>%
    addProviderTiles(providers$OpenStreetMap) %>%
    addPolygons(
      fillColor = ~pal(total_event_days),
      weight = 1,
      opacity = 1,
      color = "#888",
      dashArray = "3",
      fillOpacity = 0.7,
      highlightOptions = highlightOptions(
        weight = 3,
        color = "#555",
        dashArray = "",
        fillOpacity = 0.8,
        bringToFront = TRUE
      ),
      label = ~paste0("Number of events: ", as.character(total_event_days),
                      "\n   [FIPS Code: ", as.character(GEOID), "]"),
      labelOptions = labelOptions(
        style = list("font-weight" = "normal", padding = "3px 8px"),
        textsize = "15px",
        direction = "auto"
      )
    ) %>%
    addLegend("bottomright", pal = pal, values = ~total_event_days,
              title = "Total Number of Days Impacted by Extreme Events")
}

filtered_data <- events_frequency_counties %>% 
                  filter(event_type == "Extreme Heat Event", 
                         year_numerical == 2022)
class(filtered_data)

create_leaflet_map(filtered_data, event_type = "Extreme Heat Event")
```

## GridEX Shiny app demo

<https://ehce.connect.hms.harvard.edu/us_ehe_ece>

I)  Explore extreme events dynamics daily impacts
II) Explore extreme events aggregate yearly impacts
