---
title: "PAA2024 - Extreme Heat and Cold Events Exposure Assessment with GridEX"
subtitle: "Exploring the cumulative impacts of extreme events"
author: "Pedram Fard, Researcher at Harvard Medical School"
email: pedram_fard@hms.harvard.edu

format: html
editor: visual
---

email: pedram_fard\@hms.harvard.edu

### April 2024, PAA Annual Meeting

## Session 2 - Part D - Exploring the cumulative impacts of extreme events

Author: Pedram Fard, Researcher at Harvard Medical School\
email: pedram_fard\@hms.harvard.edu

In this notebook we work with the vector layers representing the cumulative impacts of extreme heat and cold events (EHE/ECEs) in the US.

For more information see: [Spatio-temporal interpolation and delineation of extreme heat events in California between 2017 and 2021](https://www.sciencedirect.com/science/article/pii/S0013935123017887?dgcid=rss_sd_all)

```{r, message=TRUE, warning=FALSE, error=TRUE}

library(sf)
library(dplyr)
library(leaflet)

#workshop_dataset <- "../Data"
## Check out the data folders for today workshop
#list.files(workshop_dataset, full.names = TRUE)

# Load the spatial layers of administrative boundaries 
census_data_dir <- "../Data/Census_admin_boundaries"

us_states <- st_read(file.path(census_data_dir, "tigris_states_geo_CRS5070.gpkg"))
us_counties <- st_read(file.path(census_data_dir, "tigris_counties_geo_CRS5070.gpkg"))
```

"S190" indicates that we are loading the County level data. "T102" means this file contains aggregate yearly frequency of extreme event days.

```{r}
events_frequency_counties <- readRDS(file.path("../Data/GridEX_C15_T102_by_type_by_year_summaries", "C15_S190_T102_events_by_type_by_year_Counties.rds"))

glimpse(events_frequency_counties)

```

## GridEX Shiny app demo

https://ehce.connect.hms.harvard.edu/us_ehe_ece/

I)  Explore extreme events dynamics and aggregate impacts
II) Explore extreme events aggregate impacts
