---
title: "PAA2024 - Extreme Heat and Cold Events Exposure Assessment with GridEX"
subtitle: "Exploring the cumulative impacts of extreme events"
author: "Pedram Fard, Researcher at Harvard Medical School"
email: pedram_fard@hms.harvard.edu

format: html
editor: visual
---

email: pedram_fard\@hms.harvard.edu

### April 2024, PAA Annual Meeting

## Session 2 - Part D - Exploring the cumulative impacts of extreme events

In this notebook we work with the raster layers representing the cumulative impacts of extreme heat and cold events (EHE/ECEs) in the US.

For more information see: [Spatio-temporal interpolation and delineation of extreme heat events in California between 2017 and 2021](https://www.sciencedirect.com/science/article/pii/S0013935123017887?dgcid=rss_sd_all)

```{r, message=TRUE, warning=FALSE, error=TRUE}
library(sf)
library(terra)

#workshop_dataset <- "../Data"
## Check out the data folders for today workshop
#list.files(workshop_dataset, full.names = TRUE)

# Load the spatial layers of administrative boundaries 
census_data_dir <- "../Data/Census_admin_boundaries"

us_states <- st_read(file.path(census_data_dir, "tigris_states_geo_CRS5070.gpkg"))
us_counties <- st_read(file.path(census_data_dir, "tigris_counties_geo_CRS5070.gpkg"))
```

```{r}
daily_raster_dir <- "../Data/GridEX_T002_daily_raster_dataset"
list.files(daily_raster_dir)
```

### Explore the stacked dataset of ECE for a sample year (2022)

"S422" tells us that we're bringing in gridded data covering areas experienced extreme cold event. "T002" tells us that this file has daily temporal resolution.

```{r}
ece_daily_stack_2022 <- rast(file.path(daily_raster_dir, "Y2022_S422_T002_ece_dissolved_grids_500_stack.tif"))

print(ece_daily_stack_2022)
```

```{r}
plot(ece_daily_stack_2022)
```

### Explore the cumulative impact of ECEs for a sample year (2022)

```{r}
yearly_raster_dir <- "../Data/GridEX_S420_T902_cumulative_raster_dataset"
list.files(yearly_raster_dir)
```

"S422": raster layer presenting extreme cold event

"T102": the cell values represent the total number of events in one year

```{r}
total_ece_days_2022 <- rast(file.path(yearly_raster_dir, "Y2022_S422_T102_ece_dissolved_grids_500_cumulative.tif"))

print(total_ece_days_2022)

```

```{r}
minmax(total_ece_days_2022)
#minmax(total_ece_days_2022)[2,1]

plot(total_ece_days_2022)

plot(us_states[2], add = T, col = NA)
```

"S422": raster layer presenting extreme cold event

"T102": the cell values represent the total number of events in one year

"C15\_...\_stack": The temporal coverage is the entire 15 years (2008-2002)

```{r}
total_ece_days_stack <- rast(file.path(yearly_raster_dir, "C15_S422_T102_ece_unified_grids_500_stack.tif"))

print(total_ece_days_stack)
```

```{r}
plot(total_ece_days_stack)
plot(us_states[2], add = T, col = NA)
```

### Explore the cumulative impact of ECEs for the entire period

"S422": raster layer presenting extreme cold event

"T902": the cell values represent the total number of events in the entire 15 years

```{r}
cumulative_total_ece_days <- rast(file.path(yearly_raster_dir, "C15_S422_T902_ece_unified_grids_500_cumulative.tif"))

print(cumulative_total_ece_days)
```

### Total number of ECE days between 2008 and 2022

```{r}
#plot(cumulative_total_ece_days)
#plot(us_states[2], add = T, col = NA)

# Define a color ramp, here's an example using a blue to red gradient
color_ramp <- colorRampPalette(c("purple", "navy", "blue", "cyan2", "lightblue"))

# Define the number of colors you want to use
num_colors <- 100

# Create the colors using the defined color ramp, but reversed with rev()
colors <- rev(color_ramp(num_colors))

# Plot the SpatRaster with the reversed colors
plot(cumulative_total_ece_days, col=colors)
plot(us_states[2], add = T, col = NA)
```

```{r}
minmax(cumulative_total_ece_days)
```

```{r}
hist(cumulative_total_ece_days)

## Extract the values from the raster
#values <- values(cumulative_total_ece_days)

## Remove NA values if necessary
#values <- na.omit(values)

## Use the base R hist function for more control over the histogram
#hist(values, breaks=30, main="Raster Value Distribution")
```
